---
title: "R Code"
output: pdf_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is a report of the work that has been done by Group K for the Janta ka Mood Intern.
The members were Deepak, Abhijeet and Kartikeya.

```{r data, message=FALSE}
library(tidyverse)

data <- readxl::read_xlsx("C:\\Users\\hp\\Downloads\\mp_ac_152_split.xlsx") # the electoral roll for the 152 constituency
data <- data[!duplicated(data$VoterID),]
form20 <- as.data.frame(readxl::read_xlsx("C:\\Users\\hp\\Downloads\\152_Form20.xlsx")) # the form20 data

```

We first calculate the vote share of both the candidates.

```{r vote share}
form20["Vote Share INC"] <- form20$Digvijay_Singh.INC.*100/form20$Total_Votes
form20["Vote Share BJP"] <- form20$Pragya_Thakur.BJP.*100/form20$Total_Votes
head(form20)
```

We then calculate the average age of each polling booth.

```{r age}
avg <- aggregate(data$Age, list(data$PartNo), FUN=mean)
head(avg)
```

This is a plot of the frequency of the average age distribution in the overall constituency no. 152. As we can see the
average age groups more or less lie within 

```{r histogram, echo=FALSE}
hist(avg$x, xlab = "Age", main = "Age distribution", col = c("blue", "red", "purple", "green", "violet", "orange", 
                                                            "yellow"), breaks = 50)
```

We then ensure that the same polling booths are in each dataset and then sort the sheets with their counts and plot the polling booth wise distribution and run a regression line to give an idea about the overall effect of age on the vote share.

```{r age vote share}
final <- avg %>% arrange(x)
form20_a <- form20[2:282,]
plot(final$x, form20_a$`Vote Share INC`, type = "l", xlab = "Average Age", 
     ylab = "Vote Share", main = "Age - Combined")
abline(lm(form20_a$`Vote Share INC`~final$x), col = "red")
lines(final$x, form20_a$`Vote Share BJP`, col="blue")
abline(lm(form20_a$`Vote Share BJP`~final$x), col = "purple")
legend("topleft", legend = c("INC", " ", "BJP", " "), col = c("black", "red", "blue", "purple"), 
       lty = 1, bg = "lightblue")
```

We can infer from this plot that as the average age increases, the vote share for the BJP decreases and that for the INC increases. This means that voters with a higher number of elder voters prefer to vote for the Congress more than they prefer to vote for the BJP.

We move on to the gender analysis from the age analysis. We group the data by polling booths and genders to get a count of the genders in each polling booth. The third gender is being ignored for the time being.

```{r gender}
sex <- as.data.frame(data %>% group_by(PartNo, Sex) %>% count())
head(sex)
```

We separate the male and female numbers from the counts calculated above and thereafter calculate the perccentage of male and female populations respectively.

```{r gender percent calculation}
female <- sex[which(sex$Sex=="F"),]
male <- sex[which(sex$Sex=="M"),]
male["Percentage"] <- male$n/(male$n+female$n)
female["Percentage"] <- female$n/(male$n+female$n)
```

This is a plot of the frequency of the percentage of male and female population in the overall constituency no. 152. 

```{r gender percentage, echo=FALSE}
hist(male$Percentage,breaks = 50, xlab = "% Male Population", main = "Male Population distribution - 152", col = c("blue", "red", "purple", "violet"))

hist(female$Percentage,breaks = 50, xlab = "% Female Population", main = "Female Population distribution - 152", col = c("blue", "red", "purple", "violet"))
```

We filter the outliers and We plot the vote shares of the parties against the % male population and the % female population respectively.

```{r male plot}
male <- male %>% arrange(Percentage) %>% filter(Percentage < 0.60)
form20_m <- form20[form20$PartNo %in% male$PartNo, ]
plot(male$Percentage, form20_m$`Vote Share INC`, type = 'l', main = "Male Votes Combined", xlab = "% Male Population")
abline(lm(form20_m$`Vote Share INC`~male$Percentage), col = "red")
lines(male$Percentage, form20_m$`Vote Share BJP`, col = "blue")
abline(lm(form20_m$`Vote Share BJP`~male$Percentage), col = "purple")
legend("topleft", legend = c("INC", " ", "BJP", " "), col = c("black", "red", "blue", "purple"), 
       lty = 1, bg = "lightblue")

```

```{r female plot}
female <- female %>% arrange(Percentage) %>% filter(Percentage > 0.40)
form20_f <- form20[form20$PartNo %in% female$PartNo, ]
plot(female$Percentage, form20_f$`Vote Share INC`, type = 'l', main = "Female votes Combined", xlab = "% Female Population")
abline(lm(form20_f$`Vote Share INC`~female$Percentage), col = "red")
lines(female$Percentage, form20_f$`Vote Share BJP`, col = "blue")
abline(lm(form20_f$`Vote Share BJP`~female$Percentage), col = "purple")
legend("topleft", legend = c("INC", " ", "BJP", " "), col = c("black", "red", "blue", "purple"), 
       lty = 1, bg = "lightblue")

```

We scraped the internet to collect muslim first and last names in order to classify the population in the electoral on the basis of either Muslim or Hindu. The assumptions made is that the percentage of minorities that are not Hindu or Muslim is low and that of the population is not Muslim then it is bound to be Hindu.

We had pre-divided the names in the electoral roll into first, middle and last names. We first compare the last names to the dataset for muslim last names and isolate that data from the dataset.
We then follow the same process with the remaining data set for the muslim first names to ensure the remaining names are also classified.

```{r religion classification}
data2 <- readxl::read_xlsx("C:\\Users\\hp\\Downloads\\Religion Last Names.xlsx", sheet = 1)
data3 <- readxl::read_xlsx("C:\\Users\\hp\\Downloads\\Religion Last Names.xlsx", sheet = 2)

lol1 <- data[tolower(data$LastNameEng) %in% tolower(data2$Name),]
lol1["Religion"] <- "muslim"
dataf <- subset(data, !(tolower(data$LastNameEng) %in% tolower(data2$Name)))
lol2 <- dataf[tolower(dataf$FirstNameEng) %in% tolower(data3$Name),]
lol2["Religion"] <- "muslim"
dataf <- subset(dataf, !(tolower(dataf$FirstNameEng) %in% tolower(data3$Name)))
dataf["Religion"] <- "hindu"
dat <- rbind(lol1, lol2, dataf)
```

We group the combined dataset on the basis of part no. and the religion and then get the count of the population in each. We further separate the percentages of each population.

```{r religion counts}
lmao <- as.data.frame(dat %>% group_by(PartNo, Religion) %>% count())
hindu <- lmao[which(lmao$Religion=="hindu"),]
muslim <- lmao[which(lmao$Religion=="muslim"),]

#giving values to booths in which the count is equal to 0

booth_38 <- c(38, "muslim", 0)
booth_87 <- c(87, "muslim", 0)
muslim <- rbind(muslim[1:37, ], booth_38, muslim[38:279, ])
muslim <- rbind(muslim[1:86, ], booth_87, muslim[87:280, ])
hindu$n <- as.numeric(hindu$n)
muslim$n <- as.numeric(muslim$n)

#Calculating the percentage of hindu and muslim population

hindu["Percentage"] <- hindu$n/(hindu$n + muslim$n)
muslim["Percentage"] <- muslim$n/(hindu$n + muslim$n)
```

These are the frequency distibutions of the Hindu and Muslim population in the 152 constituency

```{r plots, echo=FALSE}
hist(hindu$Percentage,breaks = 50, xlab = "% hindu Population", main = "Hindu population distribution - 152", 
     col = c("blue", "red", "purple", "violet"))

hist(muslim$Percentage,breaks = 50, xlab = "% muslim Population", main = "Muslim population distribution - 152",
     col = c("blue", "red", "purple", "violet"))

```

We then plot the vote share of each polling booth against the percentage of Hindu and Muslim Populations and then draw a regression line to get an overall idea of a voting relation between the parameters.

```{r hindu plots}

hindu <- hindu %>% arrange(Percentage) %>% filter(Percentage > 0.8) #arranging in order
form20_h <- form20[form20$PartNo %in% hindu$PartNo, ]
plot(hindu$Percentage, form20_h$`Vote Share INC`, ylim = c(0,100), type = 'l', main = "Hindu Votes - Combined", xlab = "% Hindu Population")
abline(lm(form20_h$`Vote Share INC`~hindu$Percentage), col = "red")
lines(hindu$Percentage, form20_h$`Vote Share BJP`, col = "blue")
abline(lm(form20_h$`Vote Share BJP`~hindu$Percentage), col = "purple")
legend("topleft", legend = c("INC", " ", "BJP", " "), col = c("black", "red", "blue", "purple"), 
       lty = 1, bg = "lightblue")
```
It is astounding to see that even though the average vote shares are much higher, the regression line for the INC increases as the % hindu population increases contrary to popular opinion.

```{r muslim plots}
muslim <- muslim %>% arrange(Percentage) %>% filter(Percentage < 0.2) #arranging in order
form20_m <- form20[form20$PartNo %in% muslim$PartNo, ]
plot(muslim$Percentage, form20_m$`Vote Share INC`, ylim = c(0,100), type = 'l', main = "Muslim Votes - Combined", xlab = "% Muslim Population")
abline(lm(form20_m$`Vote Share INC`~muslim$Percentage), col = "red")
lines(muslim$Percentage, form20_m$`Vote Share BJP`, col = "blue")
abline(lm(form20_m$`Vote Share BJP`~muslim$Percentage), col = "purple")
legend("topleft", legend = c("INC", " ", "BJP", " "), col = c("black", "red", "blue", "purple"), 
       lty = 1, bg = "lightblue")
```
Owing to the fact that the percentage of muslim population in 152 constituency is very less in most of the polling booths, there presence doesn't really make much of a difference as can be seen by the regression lines which have ~0 slope. However a slight decrease can be seen in the vote share of the BJP and a slight increase in that of the INC.

We then proceeded to create our R shiny dashboard. Depending on the candidate you can make a choice between Digvijay Singh and Sadhvi Pragya Thakur, the analysis changes accordingly.

Clicking the bars of the age groups leads you to the voters of that age group specifically.

We can see the percentage of voters of each demographic. as well as the vote distribution. Depending on the candidate when we select the to include the other votes, how the voting pattern would have changed can be seen in the pie chart presented.

The colour of the winning margin boxes change to specify the winning party as well

Some preliminary data manipulation has been done for ease of working

```{r shiny app, message=FALSE}
data <- as_data_frame(dat)
data <- data %>% drop_na(PollingStationAddressEn)
data$Age_Group <- ifelse(data$Age < 28, "18-27", ifelse(data$Age < 38, "28-37", ifelse(data$Age < 48, "38-47", ifelse(data$Age < 58, "48-57", ifelse(data$Age < 68, "58-67", "68+"))))) # setting the age groups
colnames(data)
d <- dim(data)

const <- as.list(c("Digvijay Singh", "Sadhvi Pragya Thakur"))

#Formatting the data

data$PartNo <- sprintf("%03d", as.numeric(data$PartNo))
form20$PartNo <- sprintf("%03d", as.numeric((form20$PartNo)))
head(form20$PartNo)
y <- as_tibble(cbind("000 Entire Constituency", d[1]))
colnames(y) <- c("Booth", "n")
y


data["Booth"] <- paste(data$PartNo, data$PollingStationAddressEn) #Naming the data
lol <- data %>% group_by(Booth) %>% count()
lol <- rbind(y , lol)
lol %>% drop_na(Booth)
lol <- cbind(lol, form20)
lol["Voting Percentage"] <- (as.numeric(lol$Total_Votes)/as.numeric(lol$n))
lol["Voting Difference"] <- abs(as.numeric(lol$Digvijay_Singh.INC.) - as.numeric(lol$Pragya_Thakur.BJP.))
lol["Non Votes"] <- as.numeric(lol$Total_Votes) - (as.numeric(lol$Digvijay_Singh.INC.) + as.numeric(lol$Pragya_Thakur.BJP.))
lol$Winner <- ifelse(lol$Digvijay_Singh.INC. < lol$Pragya_Thakur.BJP., "BJP", "INC")
head(lol)

#number of swing votes

hehe <- data %>%
  group_by(PartNo)%>%
  filter(Age <= 23) %>%
  tally() %>%
  pull() %>%
  as.data.frame()

huhu <- data %>%
  filter(Age <= 23) %>%
  tally() %>%
  pull() %>%
  as.data.frame()

f <- rbind(huhu, hehe)

lol["Swing"] <- f

lol["Other"] <- lol$Total_Votes - lol$Digvijay_Singh.INC. - lol$Pragya_Thakur.BJP.

foo1 <- cbind(as.data.frame(lol$Booth), lol$Digvijay_Singh.INC., lol$Pragya_Thakur.BJP. + lol$Other)
colnames(foo1) <- c("Booth", "INC - Digvijay Singh", "BJP - Sadhvi Pragya Thakur")
foo2 <- cbind(as.data.frame(lol$Booth), lol$Digvijay_Singh.INC. + lol$Other, lol$Pragya_Thakur.BJP.)
colnames(foo2) <- c("Booth", "INC - Digvijay Singh", "BJP - Sadhvi Pragya Thakur")
foo3 <- cbind(as.data.frame(lol$Booth), lol$Digvijay_Singh.INC., lol$Pragya_Thakur.BJP., lol$Other)
colnames(foo3) <- c("Booth", "INC - Digvijay Singh", "BJP - Sadhvi Pragya Thakur", "Others")

library(shiny)
library(shinydashboard)
library(ECharts2Shiny)
library(scales)
library(r2d3)

ui <- dashboardPage(
  dashboardHeader(
    title = "Polling Booth Dashboard",
    titleWidth = 200
  ),
  dashboardSidebar(
    width = 300,
    selectInput(
      inputId = "party",
      label = "Select Candidate:",
      choices = const,
      selectize = FALSE
    ),
    sidebarMenu(
      selectInput(
        inputId = "booth",
        label = "Polling Booth:",
        choices = lol[1],
        selected = 1,
        selectize = TRUE
      ),
      actionLink("remove", "Remove detail tabs")
    )
  ),
  dashboardBody(
    tags$head(tags$script(loadEChartsLibrary())),
    tabsetPanel(
      id = "tabs",
      tabPanel(
        loadEChartsLibrary(),
        title = "Main Dashboard",
        value = "page1",
        fluidRow(
          valueBoxOutput("total_voters", width = 2),
          valueBoxOutput("total_votes", width = 2),
          valueBoxOutput("Vot_per", width = 2),
          valueBoxOutput("Vot_diff", width = 2),
          valueBoxOutput("Vot_non", width = 2),
          valueBoxOutput("total_swing", width = 2),
        ),
        fluidRow(
          valueBoxOutput("male_vote", width = 3),
          valueBoxOutput("female_vote", width = 3),
          valueBoxOutput("hindu_vote", width = 3),
          valueBoxOutput("muslim_vote", width = 3),
        ),
        fluidRow(),
        fluidRow(
          column(
            width = 4,
            d3Output("age_groups")
          ),
          column(
            width = 3,
            selectInput(inputId = "select", 
                        label = h3("Select Category"),
                        choices = list("Current Vote Share" = 3, "Vote Share with others included" = 2),
                        selected = 3),
            valueBoxOutput("text", width = 100),
          ),
          column(
            width = 5,
            tags$div(id="test_2", style="width:100%; height:300px;"),  # Specify the div for the chart. Can also be considered as a space holder
            deliverChart(div_id = "test_2"),
          )# Deliver the plotting
        )
      )
    )
  )
)

server <- function(input, output, session) {
  tab_list <- NULL
  
  base_voters <- reactive({
    res <- data
    if(input$booth != "000 Entire Constituency") res <- filter(res, Booth ==input$booth)
    res
  })
  base_votes <- reactive({
    res1 <- lol
    res1 <- filter(res1, Booth ==input$booth)
    res1
  })
  share_votes <- reactive({
    res2 <- foo1
    res3 <- foo2
    res4 <- foo3
    if(input$select==2 && input$party== "Sadhvi Pragya Thakur"){
      rest <- filter(res2, Booth==input$booth)
      res <- data.frame(
        name = c("INC - Digvijay Singh", "BJP - Sadhvi Pragya Thakur"),
        value = c(rest$`INC - Digvijay Singh`, rest$`BJP - Sadhvi Pragya Thakur`)
      )
    }
    if(input$select==2 && input$party == "Digvijay Singh"){
      rest <- filter(res3, Booth==input$booth)
      res <- data.frame(
        name = c("INC - Digvijay Singh", "BJP - Sadhvi Pragya Thakur"),
        value = c(rest$`INC - Digvijay Singh`, rest$`BJP - Sadhvi Pragya Thakur`)
      )
    }
    if(input$select==3){
      rest <- filter(res4, Booth==input$booth)
      res <- data.frame(
        name = c("INC - Digvijay Singh", "BJP - Sadhvi Pragya Thakur", "Others"),
        value = c(rest$`INC - Digvijay Singh`, rest$`BJP - Sadhvi Pragya Thakur`, rest$Others )
      )
    }
    res
  })
  output$total_voters <- renderValueBox({
    base_voters() %>%
      tally() %>%
      pull() %>%
      as.integer() %>%
      prettyNum(big.mark = ",") %>%
      valueBox(subtitle = "Number of Voters")
  })
  output$total_swing <- renderValueBox({
    base_voters() %>%
      filter(Age <= 23) %>%
      tally() %>%
      pull() %>%
      as.integer() %>%
      prettyNum(big.mark = ",") %>%
      valueBox(subtitle = "Number of Swing Voters")
  })
  output$total_votes <- renderValueBox({
    base_votes() %>%
      select("Total_Votes") %>%
      pull() %>%
      as.integer() %>%
      prettyNum(big.mark = ",") %>%
      valueBox(subtitle = "Number of Votes Given", col = "blue")
  })
  output$Vot_per <- renderValueBox({
    base_votes () %>%
      select(("Voting Percentage")) %>%
      as.numeric () %>%
      percent(accuracy = 0.01, decimal.mark = ".") %>%
      valueBox(subtitle = "Voting Percentage", col = "blue")
  })
  output$Vot_diff <- renderValueBox({
    haha <- base_votes()%>%select("Winner")
    if(haha == "BJP"){
      base_votes () %>%
        select(("Voting Difference")) %>%
        as.integer () %>%
        prettyNum(big.mark = ",") %>%
        valueBox(subtitle = "Winning Margin - Sadhvi Pragya", col = "orange") 
    }else if(haha == "INC"){
      base_votes () %>%
        select(("Voting Difference")) %>%
        as.integer () %>%
        prettyNum(big.mark = ",") %>%
        valueBox(subtitle = "Winning Margin - Digvijay Singh", col = "green") 
    }
  })
  output$Vot_non <- renderValueBox({
    base_votes () %>%
      select("Non Votes") %>%
      as.integer () %>%
      prettyNum(big.mark = ",") %>%
      valueBox(subtitle = "Non BJP-INC Votes", col = "blue")
  })
  output$age_groups <- renderD3({
    res <- base_voters() %>%
      group_by(Age_Group) %>%
      tally() %>%
      collect() %>%
      mutate(
        y = n,
        x = Age_Group
      ) %>%
      select(x, y)
    
    res <- res %>% mutate(label = x)
    r2d3(res, "col_plot.js")
  })
  
  get_details <- function(group = NULL, religion = NULL) {
    # Create a generic details function that can be called
    # by different dashboard events
    res <- base_voters()
    if (!is.null(group)) res <- filter(res, Age_Group == group)
    if (!is.null(religion)) res <- filter(res, Religion == religion)
    
    res %>%
      select(
        Booth, HouseNoEn, VoterNameEn, Sex, Age, VoterID, Religion
      ) %>%
      group_by(HouseNoEn)%>%
      collect() %>%
      arrange(VoterNameEn, .by_group = TRUE)
  }
  observeEvent(input$column_clicked != "", {
    
    age_g <- input$column_clicked
    tab_title <- paste(
      input$PartNo, "- Age Group", age_g
    )
    if (!(tab_title %in% tab_list)) {
      appendTab(
        inputId = "tabs",
        tabPanel(
          tab_title,
          DT::renderDataTable(
            get_details(group = age_g)
          )
        )
      )
      tab_list <<- c(tab_list, tab_title)
    }
    updateTabsetPanel(session, "tabs", selected = tab_title)
  },
  ignoreInit = TRUE
  )
  output$male_vote <- renderValueBox({
    base_voters() %>%
      filter(Sex == "M") %>%
      tally() %>%
      pull() %>%
      as.integer() %>%
      prettyNum(big.mark = ",") %>%
      valueBox(subtitle = "Number of Male Voters")
    
  })
  output$female_vote <- renderValueBox({
    base_voters() %>%
      filter(Sex == "F") %>%
      tally() %>%
      pull() %>%
      as.integer() %>%
      prettyNum(big.mark = ",") %>%
      valueBox(subtitle = "Number of Female Voters")
    
  })
  output$hindu_vote <- renderValueBox({
    base_voters() %>%
      filter(Religion == "hindu") %>%
      tally() %>%
      pull() %>%
      as.integer() %>%
      prettyNum(big.mark = ",") %>%
      valueBox(subtitle = "Number of Hindu Voters")
    
  })
  output$muslim_vote <- renderValueBox({
    base_voters() %>%
      filter(Religion == "muslim") %>%
      tally() %>%
      pull() %>%
      as.integer() %>%
      prettyNum(big.mark = ",") %>%
      valueBox(subtitle = "Number of Muslim Voters")
    
  })
  output$text <- renderValueBox({
    res <- share_votes()
    out <- ifelse(res[1,2]>res[2,2], "INC Wins", "BJP Wins")
    out <- out %>%
      as.character() %>%
      valueBox(subtitle = "Who wins?")
      
  })
  observeEvent(share_votes(),
               {
                 renderPieChart(div_id = "test_2",
                                data = share_votes(),
                                radius = "70%",center_x = "50%", center_y = "50%")
               })
  observeEvent(input$remove, {
    # Use purrr's walk command to cycle through each
    # panel tabs and remove them
    tab_list %>%
      walk(~ removeTab("tabs", .x))
    tab_list <<- NULL
  })
}

shinyApp(ui, server)

```